Explorations to Optimize the Parareal Algorithm \\to Solve ODEs in Given Applications

### 2014-12-10 paper.tex
pushing lease no conflicts

\begin{equation}
\textnormal{Eff}_{\textnormal{strong}} = \frac{T(1)}{n \times T(n)} \times 100\%
\end{equation}

@@

\begin{equation}
\textnormal{Eff}_{\textnormal{weak}} = \frac{T(1)}{T(n)} \times 100\%
\end{equation}

@@

\begin{equation}
t + k \left(t + \frac{Qt}{N} \right).
\end{equation}

@@

For the parallelism by space, we are dividing our grid on which we which to solve our wave diffusion equation. This division is inherently parallelizable and allows us to follow a very similar approach in computing this as in the serial version - just divided in space, where the stencil approach is well-defined.

@@

To delve one step deeper in our analysis, parallel communications are usually expressed by architecture parameters - $\alpha$, $\beta$ and $\gamma$

@@

The Parareal method TODO does this scale well?

@@

\subsection{Explored Tests}

@@

\begin{itemize}
\item
\item
\end{itemize}

@@

We implemented the Parareal algorithm in Python, using mpi4py to parallelize it.
Our coarse operator was forward Euler step, with 100 steps, while the fine
operator was forward Euler, with $Q \cdot 100$ steps, where $Q$ is the quality
factor. We tested the Parareal algorithm on two sets of differential equations:

>>>>

\begin{equation}
\textnormal{Eff}_{\textnormal{strong}} = \frac{T(1)}{n * T(n)} * 100\%
\end{equation}

@@

\begin{equation}
\textnormal{Eff}_{\textnormal{weak}} = \frac{T(1)}{T(n)} * 100\%
\end{equation}

@@

\begin{equation}
t + k \left(t + \frac{Qt}{N} \right).
\label{eq:scheme}
\end{equation}

@@

For the parallelism by space, we are dividing our grid on which we which to solve our wave diffusion equation. This division is inherently parallelizable and allows us to follow a very similar approach in computing this as in the serial version - just divided in space, where the stencil approach is well-defined. A fully parallelizable approach allows for $n$ times speedup and no computational tradeoffs if we ignore communication. As we are dividing our processors into a processor grid (dividing a 2D space into a 2D arrangement of processors) we are only limited by when the processor grid size approaches the problem size - upon which further processors would not receive any tasks since that would require a division by less than a unit of computation.

@@

To delve one step deeper in our analysis, parallel communications are usually expressed by architecture parameters - $\alpha$, $\beta$ and $gamma$ explained in the below table.

\begin{table}
\centering
\begin{tabular}{|c|l|}
{\bf Variable} & {\bf Description}\\
\hline\hline
$\alpha$ & Latency of communication. (sec)\\
$\beta$ & Inverse communication bandwidth. (sec/byte)\\
$\gamma$ & Inverse computational performance. (sec/flop)\\
\hline
\end{tabular}
\caption{Table of Communication Architecture Parameters}
\label{tab:notation}
\end{table}

\subsubsection{Parareal}
TODO THIS IS CONFUSING Wesley

\subsubsection{Parallel By Space}
TODO THIS IS CONFUSING TOO Wesley

@@

We first try to express $B$ as function of the quality factor $Q$ and the number of correction steps of the parareal $k$.

From our analysis in equation~\ref{eq:scheme}, when we set $N$ to 1 for the generic setup, we see that our total computation is $t + K(t+ Qt)$.  Of this, the only parallelizable part is the kQt term.  Thus to find the expression of the parallelizable portion, we simplify:

\[

\]

Now to replace this into Amdahl's Law, we have:

@@

\subsection{Planned Tests to Run}

@@

\begin{itemize}
\item General Scaling with Incrased $N$
\item Adjusting $K$ in the Parareal
\item Using Different Integration Schemes in Parareal for $g_{\textnormal{fine}}$
\item Run the Solver on Various Complexities of ODEs
\end{itemize}

@@

We implemented the Parareal algorithm in Python, using mpi4py to parallelize it.
Our coarse operator was forward Euler step, with 100 steps, while the fine
operator was forward Euler, with $Q * 100$ steps, where $Q$ is the quality
factor. We tested the Parareal algorithm on two sets of differential equations:
